EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-30 draft.tex
Moves figure closer to text
Moving some stuff around
Added conclusion
This clarifies the main selling points of the paper, closes #1
Also changed title, and checked bibliography, not much more to add,
closes #2

\title{An Event-based Architecture for Multi-population Optimization Algorithms}

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a multi-container Docker application that reactively processes isolated and
heterogeneous populations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
% justification: - Mario
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

@@

To test the heterogeneous multi-population capabilities,  we compare the
performance between a homogeneous and an ensemble of multi-populations, using
Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
implementation, we used the DEAP library \cite{fortin2012deap} and for the PSO,
the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
open-source. In EvoSwarm these two algorithms are implemented as stateless
functions. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.

% Maybe this should be mentioned earlier, in the workflow? - JJ

@@

We present the results of the experiments in this section.
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%
\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{sp_size_worker}
    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
    \label{fig:spworker}
\end{figure}
%
%
% \subsection{Experiment 1}
% \label{sec:exp1results}
%
with results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

@@

>>>>

\title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a multi-container Docker application that reactively processes isolated and
heterogeneous populations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

@@

To test the heterogeneous multi-population capabilities,  we compare the
performance between a homogeneous and an ensemble of multi-populations, using
Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
implementation, we used the DEAP library \cite{fortin2012deap} and for the PSO,
the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
open-source. In EvoSwarm these two algorithms are implemented as stateless
functions.

@@

We present the results of the experiments in this section.
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%
\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{sp_size_worker}
    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
    \label{fig:spworker}
\end{figure}
%
%
% \subsection{Experiment 1}
% \label{sec:exp1results}
%
with results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

@@

In this paper we have presented the EvoSwarm framework that implements
an event-based, cloud native architecture suitable for hosting any
kind of multi-population optimization algorithm. The framework is
based in industry-standard development tools such as Docker and Docker
Compose, making easy to develop and deploy multi-algorithm distributed
experiments.

The set up of this kind of algorithms will have a big influence on the
results, since diversity is important for this kind of algorithms and
its creation and destruction play a big part of the algorithm's
performance. One of the main levers to control it in EvoSwarm is via
the number of workers that will be allocated from the beginning, which
is why we have performed an experiment mainly intended as a proof of
concept, but also devoted to show how two parameters in the framework,
number of workers and number of populations used to initially seed the
system, interact with each other. The conclusion of this experiment is
that there is an interesting interplay, with the most benefit being
obtained with a number of populations equal to the number of workers
or slightly inferior or superior, which brings algorithmic benefit
(experiment taking less time because it needs less evaluations) and
also performance benefits (experiment taking less time because
evaluations are being split between different workers.

As a conclusion, EvoSwarm is a framework designed and tested according
to industry standars which is able to bring many benefits to the
bioinspired algorithm community, by showing the possibility of
creating new, high performance, mixed algorithms, which can also be
easily deployed to the cloud, or locally to systems that simply need
to have Docker and Docker Compose installed.

It also opens new possibilities. From the dessign point of view, it
would be interesting to try and use automatic scaling, instead of
setting the number of workers by hand. This would need a certain
amount of redesign, the most important of which would be using
container orchestration systems such as Kubernetes or Docker
Swarm. Instead of setting a fixed number of workers, only the maximum
amount would need to be established, which could be done directly or
by setting an evaluation budget.

From the algorithmic point of view, there are many possible lines of
research. Mixing population-based algorithms with other algorithms
would be a possibility, but also using different instances of
population based algorithms, such as checking how Estimation of
Distribution Algorithms would work together with evolutionary
algorithms. EvoloPy also includes other population based algorithms,
which could be tested, trying to find out which sets of algorithms are
a better match to each other. All these will be explored as future
lines of work.
